
\section{Calculations}
\label{sec:calc}

\begin{frame}{Calculations}
  \begin{itemize}
  \item Performed on Apple M2.
  \item Used {\tt Python} 3.11 with {\tt python-sat} package and
    solver {\tt cadical}
  \item Heavily dependent on good cardinality encoding.
  \item Conflictor generates up to 10 counterexamples to give to
    Resolver.
  \item Confirms that $\beta_{10} = 7$ in under 5 seconds.
  \item Confirms that $\beta_{11} = 8$ in 33689 seconds.
  \item Max Sat with 3 levels symmetry breaking: 24317 seconds.
  \item Along with $\beta_{13} \le 8$ implies $\beta_{13} = 8$.
  \item Only needs to generate a small fraction of all $x$ vectors to
    show UNSAT.
  \end{itemize}
\end{frame}
\begin{frame}{Speed ups?}
  \begin{itemize}
  \item Heuristic: $\exists$ a balanced detecting matrix
    $\Leftrightarrow$ $\exists$
    one with all rows having weight $\lfloor n/2 \rfloor$.
  \item Reason: such rows detect the most balanced $x$.
  \item Proof?
  \item Using this restriction: can show $7 \times 11$ is UNSAT in 717
    seconds.
  \end{itemize}
\end{frame}
\begin{frame}{Proofs}
  \begin{itemize}
  \item Why trust a long calculation that says UNSAT?
  \item Verify the model.
  \item Many SAT solvers can emit a {\tt DRUP} proof of UNSAT.
  \item While usually long, it can be easily checked by small program.
  \item Some SMT solvers are capable of producing a \emph{Skolem
      function}: an explicit function $f(A)$ from matrices to vectors
    such that $A f(A) = 0$ for all $A$.
  \item So far have not been able to produce one for $6 \times 10$
    problem.
  \end{itemize}
\end{frame}
\begin{frame}{SMT solvers}
  \begin{itemize}
  \item Some SMT solvers can solve $\exists, \forall$ problems: {\tt
      Z3}, {\tt boolector}, {\tt bitwuzla}, {\tt YicesQS}.
  \item Encoded problem in SMT.
  \item {\tt boolector}, {\tt bitwuzla} and {\tt YicesQS} can all show
    the $6 \times 10$ problem is UNSAT in roughly 10 seconds: about
    twice as slow as custom program.
  \end{itemize}
\end{frame}
\begin{frame}{Improvements?}
  \begin{itemize}
  \item More symmetry breaking: Use group representations?
  \item Better heuristics for generating useful counterexamples.
  \item SMT: better encoding for Hamming weight.
  \end{itemize}
\end{frame}
