
\section{Symmetries}
\label{sec:symmetry}
\begin{frame}{Using Max Sat}
  \begin{itemize}
  \item Used Max Sat solver {\tt RC2}.
  \item Straightforward hitting set formulation.
  \item $\beta_n$ for $n\le 5$ in a fraction of a second.
  \item $\beta_6$ in 1.3 seconds.
  \item $\beta_7$: 15144 seconds!
  \item Problem: Lots of symmetries in the problem.
  \item Explores many equivalent paths.
  \item Hyperoctahedral group: size $2^n n!$, plus complementation.
  \end{itemize}
\end{frame}
\begin{frame}{Symmetries}
  \begin{itemize}
  \item If $A$ is a balanced detecting matrix, then so is $A'$ where
    some of the rows are complemented.
  \item Enforce: weight of each row $\le n/2$.
  \item Rows of $A$ can be permuted.
  \item Columns of $A$ can be permuted.
  \item Balanced vectors with exactly one $\pm 1$ enforce unequal
    columns.
  \item Enforce lexicographic strictly increasing rows/columns.
  \end{itemize}
\end{frame}
\begin{frame}{Symmetry Breaking}
  \begin{itemize}
  \item Many problems: there is a group acting on potential solutions.
  \item Paths in search tree that are equivalent under group must be
    explored.
  \item Solution: Try to enforce one (or small number) of equivalent
    solutions by adding predicates.
  \item Simple case: variables may be permuted: sort them lexicographically.
  \item Alternative: Generate predicates ``on the fly'' during search.
  \end{itemize}
  
\end{frame}
\begin{frame}{Double Sorting}
  \begin{itemize}
  \item Matrices where rows/columns may be permuted.
  \item Double sorting: If we alternate sorting columns/rows
    lexicographically it stabilizes (Anna Lubiw).
  \item Enforce sorting in both rows and columns.
  \item Still leaves a large number of equivalent solutions.
  \item 
  \end{itemize}
  
\end{frame}
\begin{frame}{Improving Max Sat}
  \begin{itemize}
  \item By sorting can assume first row of $A$ has contiguous 0's and
    1's.
  \item Once we have placed the first $r$ rows we can sort within
    equal columns.
  \item Can enumerate all conditions, but $r=3$ seems to work well.
  \item Brings down calculation of $\beta_7$ to a fraction of a
    second.
  \item Using $r=4$, can calculate $\beta_{10}$ in 5 seconds.
  \item But, needs a lot of time for preprocessing.
  \item Uses a lot of storage.
  \item Luks: breaking all symmetries is NP-hard.
  \item For row/column symmetries most are content to use
    double-sorting.
  \end{itemize}
\end{frame}
